---
title: "Reading data from Web"
author: "Molly Martorella"
date: "10/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
library(httr)

```

# Notes

Scraping:
download html using read_html()
extract nodes using html_nodes() and the css selector
extract content from nodes using html_text() and html_table() --> then can assign this content as a table or character vector etc.

APIs - application programming interfaces --> way to access data without needing to extract html and css etc.; sometimes API gives back weirdly formatted data rather than data in the format you would like.
Mostly accessible using http; httr contains tools for constructing http requests, use GET.

# Reading Data (using rvest):

```{r}

url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_xml <- read_html(url)

#gives all tables from website:
drug_use_xml %>%
  html_nodes(css = "table")

#want to pull out first table in the list of tables; double square brackets [[]] index a list, the dot . is the placeholder for what you piped:
table_marj <- 
  (drug_use_xml %>% html_nodes(css = "table")) %>% 
  .[[1]] %>%
  html_table()

#remove first row using slice (uses row #s rather than some criteria like filter):
table_marj <- 
  (drug_use_xml %>% html_nodes(css = "table")) %>% 
  .[[1]] %>%
  html_table() %>%
  slice(-1) %>% 
  as_tibble()

```

harry potter example:

```{r}

hpsaga_html <- 
  read_html("https://www.imdb.com/list/ls000630791/")

#use selector gadget to figure out title node. Click item you want to scrape, clikc again on things you don't want that are highlighted in yellow. Selector box gives you tag that you need which you put into the node.

title_vec <- 
  hpsaga_html %>%
  html_nodes(".lister-item-header a") %>%
  html_text()

runtime_vec <- 
  hpsaga_html %>%
  html_nodes(".runtime") %>%
  html_text()


gross_test <- 
  hpsaga_html %>%
  html_nodes(".ghost~ .text-muted+ span") %>%
  html_text()

gross_rev_vec <- 
  hpsaga_html %>%
  html_nodes(".text-small:nth-child(7) span:nth-child(5)") %>%
  html_text()

hpsaga_df <- 
  tibble(
    title = title_vec,
    rev = gross_rev_vec,
    runtime = runtime_vec)

```

# Using an API:

```{r}

nyc_water <- 
  GET("https://data.cityofnewyork.us/resource/waf7-5gvc.csv")

#content parsed as a csv file:
nyc_water <- 
  GET("https://data.cityofnewyork.us/resource/waf7-5gvc.csv") %>% 
  content("parsed")

#using json rather than api, json::lite converts to a table, then convert to a tibble:
nyc_water <- 
  GET("https://data.cityofnewyork.us/resource/waf7-5gvc.json") %>% 
  content("text") %>%
  jsonlite::fromJSON() %>%
  as_tibble()

```








